### 2020-04-16 paper/paper.tex
Updated Paper Draft
It was quite remarkable the breadth of change required from what was drafted before.

\paragraph{Abstract:}
There have been many hardware improvements in polygonal-fiber-fed,
temperature-controlled, laser-frequency-comb or etalon-calibrated,
high-resolution, extreme-precision radial-velocity spectrographs.
Have the calibration methodologies improved to match?
Here are three relevant observations:
The first is that the calibration lines or spots from an etalon or
comb (or even arc lamp) fill the spectral range with dense calibration points.
The second is that the spectrograph lives in a stabilized,
climate-controlled environment, in which the full optical system and
detectors will only vary within a tiny range of configurations or
settings.
The third is that, given this stability, every calibration image
ever taken is relevant to every science exposure ever taken; there is
no reason to calibrate every exposure independently.
The calibration methodology we propose here---\name---addresses these
three problems by going non-parametric (no more polynomials!) and then
reducing dimensionality with a robust principal-component method.
We demonstrate the success of this method with comb and arc data
from the \expres\ spectrograph.
We find... [results and so on].

\section{Introduction}

Extreme precision radial-velocity (\eprv) programs have been
outrageously successful in finding and characterizing extra-solar
planets (CITE THINGS).
These programs typically make use of spectrographs with resolutions on
the order of $10^5$, which correspond to line widths on the order of
$3000\,\mps$.
Exoplanet science happens at $3\,\mps$ precisions, and the current
instruments hope to get to $0.1\,\mps$ precisions (CITE THINGS).
This means that new spectrographs are expected to be calibrated or
stabilized to better than $10^{-4}$ of a pixel (assuming that the
spectrograph is well sampled).
And indeed, some hardware--software systems seem to be reaching these
levels (CITE THINGS).

Here we propose to simplify and improve calibration programs for
\eprv\ hardware systems with two very simple but novel ideas.
The first flows from the observation that calibration sources---which
include lamps, etalons, and laser-frequency combs (\lfc
s)---illuminate the spectrograph with dense sets of lines; almost every
location in the spectrograph image plane is surrounded by nearby,
useful calibration lines.
This recommends a calibration methodology that is
\emph{non-parametric}:
If every point in the spectrograph detector is surrounded by nearby
calibration lines, the wavelength solution can be made simply an
interpolation of the calibration data.
There is no need to choose any functional form for the wavelength
solution (such as an eighth-order polynomial, for example).
Going non-parametric will improve calibration accuracy, because
choosing a parametric form biases the calibration, and biases it
badly when the form is inappropriate (as, for example, polynomials
will be at the detector edges).

The second idea here flows from the observation that contemporary
\eprv\ instruments are incredibly stable.
Temperature-controlled, fiber-fed pectrographs vary only slightly over
the night or season, and only along a small number of axes in what you
might call ``calibration space'', or the (very high dimensional) space
of all possible wavelength solutions.
That is, not only are they stable, but they don't have many accessible
degrees of freedom.
So it is not appropriate to fit each calibration exposure or calibrate
each science exposure independently.
It makes sense to use the calibration data (or all the data) to
determine the space in which the instrument can and does vary, and
then in subsequent calibration work only permit the solution to live
in the accessible part of calibration space.
This structure is \emph{hierarchical}: The calibration data are used
not just to determine the wavlength solution, but also to determine
the possible space of wavelength solutions.
In the statistics literature this concept is often described as
\emph{de-noising}:
Every calibration exposure contains information about every other
exposure; it doesn't just constrain the instrument locally in time.
Thus every exposure can be improved (de-noised) with information
coming from every other exposure.

The method we propose here---\name---embodies these ideas.
It is a non-parametric, hierarchical, data-driven model for the
wavelength solution.
By being non-parametric, it delivers enormous freedom to the
wavelength solution to match or adapt to any instrument or detector
oddities.
By being hierarchical, it restricts that freedom tremendously, but it
does so appropriately for the observed variations in the
spectrograph.
\name\ is designed for temperature-controlled, bench-mounted, fiber-fed
spectrographs with good calibration sources, such as laser-frequency
combs, etalons, or good arc lamps.
We have in mind the \eprv\ instruments and \eprv\ science cases.
But we expect \name\ to have applications for other kinds of
spectrographs in other contexts; the motivating ideas behind this
project are true for almost every astronomical spectrograph.

\section{Method} \label{sec:method}
We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that in fact it will be living in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the spectrograph hardware.

HOGG: THIS IS NOW ALL WRONG... We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda(x,y,m,n) = f(x,y,m;\theta_{n})
\quad ,
\end{equation}
where $\theta_{n}$ is a big blob of parameters for this exposure.

A stabilized spectrograph should experience only low-degree variability, meaning the calibration of any image can be informed by the calibration of every other image.  To implement such a hierarchical model, we use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.

If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so
small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could
be replaced with a tiny model
\begin{equation}
\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
\quad ,
\end{equation}
where
$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the
spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the
fiducial calibration.
The challenge is to learn these basis functions from the data and get
the $K$ amplitudes $a_{nk}$ for every exposure $n$.

HOGG: Say that there are many ways to do this, but we use PCA for now.
Note that PCA is correct in the limit that everything is very high in SNR.

\subsection{Data} \label{sec:data}
We present results using data from \expres\, the EXtreme PRecison Spectrograph.  \expres\ is an envionmentally stabilized, fiber-fed, $R=137,000$, optical spectrograph (\lz{CJ and RB's papers}).  \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g. \lz{Wilken+ 2012, Molaro+ 2013, Probst+ 2014, from RB's paper}).  ThAr exposures are taken at the beginning and end of each night.  LFC exposures are interspersed between science exposures every ~15 minutes.  
(Discussion of number of orders covered or number of lines in LFC and/or ThAr?)

\lz{Discussion of line fitting?  Though the current line fitting causes me recurring nightmares.  I should really take the time to look into at least the peak fitting more.}

For the results presented in this paper, we work with the 1D, optimally extracted data.  This removes the y-dependence from equations 1 and 2.

\expres\ exposures are separated into different ``instrumental epochs," which correspond to changes to the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources (\lz{(worth including table that explicitly defines epochs?)}).  Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.

\subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
Let $\{exp_n\}$ be the set of calibration exposures within an instrument epoch used to construct $g_k(x,y,m)$.  We first identify the full list of lines we can expect to find in a calibration image from this epoch.  A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength of that line, $\lambda$.  Let $\{l_{m,\lambda}\}$ be the complete list of unique lines found in $\{exp_n\}$.

For each line, $l_{m,\lambda}$, we then find the fitted line center, $x_{n,l}$, in pixel space, for that line for each exposure.  If this line is missing from an exposure (e.g. it was covered by noise or off the edge of its usual order), we insert a $NaN$ for that line center instead.

Lines that are missing from greater than some percent of exposures are cut.  Exposures that are missing greater than some percent of lines are also cut (\lz{This will miss low signal exposures that suffer most in the first few orders but can recover in later orders; can't decide if this is a feature or a bug}).  This threshold percentage is 50\% for LFC exposures and 30\% for ThAr exposures.

Missing line center measurements are replaced with denoised estimates using iterative PCA.  For each missing line center measurement, we initialzie their value to the mean of the nearest $n$ exposures in time, where $n$ is typically 9 for LFC exposures and $15$ for ThAr.  We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, for which we use the mean line center value for each identified line.  We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and fiducial line center for each exposure.  This PCA gives us $a_{nk}$ and $g_k(x,y,m)$.

Using those values with equation 2 and some small integer (e.g. 2) for K, we can reconstruct a de-noised matrix of line centers for each exposure.  Missing line center measurements are then replaced with values from this reconstruction.  We do this iteratively until the values of the missing line centers change by less than 0.01\% of a pixel.  This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."

\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: \;
\While{change in bad pixels $>$ 0.01\%}{
	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
	let $a_{n,k} = U\cdot \Sigma$ and $V = g_k(x,y,m)$\;
	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ where $K=2$\;
	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ is $NaN$
	}
\caption{Hierarchical De-Noising}
\end{algorithm}

\subsection{Non-Parametric Calibration} \label{sec:nonparam}
We can interpolate $a_{nk}$ with respect to time to determine $\lambda(x,y,m,n)$ at any time.  Once we have the line center for each line, which has an associated known wavelength, we can interpolate the known wavelengths over line centers to any position in x in a given order.  For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.

We chose to implement both interpolations with a cubic spline.  Interpolation of wavelengths over line centers is done order by order.

\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: new time, new line centers, new orders\; %command not working for w/e reason
Let $t_n$ be the time associated with exposure $n$\;
let $t_n'$, $x_n'$, and $m_n'$ be the new time, line centers, and orders we want wavelengths for\;
Let $a_{n,k}'$ be the interpolation of $a_{n,k}$ to time $t_n'$\;
$\lambda(x,y,m,n)' = g_0(x,y,m) + \sum_{k=1}^K a_{nk}'\,g_k(x,y,m)$ where $K=2$\;
\For{each unique $m_n'$}{
	${l_n,m}$ from $\lambda(x,y,m,n)'$
	interpolate ${l_n,m}$ onto $x_n'$\;
	}
\caption{Non-Parametric Wavelength Solution}
\end{algorithm}
\lz{What's fancy math symbol for interpolate?}

\subsection{Choice Choices} \label{sec:choices}
(Before or after results presented in validation test and ThAr sections?)

We're still testing most of these (will use that as excuse to write this section later)
\begin{itemize}
	\item Value of K for denoising
	\item Type of interpolation (cubic spline vs nearest neighbor or linear)
	\item Interpolation order (interpolate time and x direction together?  Individually?  Which one first?)
\end{itemize}

Reason for not linear interpolation:
Due to the dispersion intrinsic to echelle spectrographs, the wavelength change between pixels is greater at greater wavelengths.  This means that the function of wavelength given pixel across an order will necessarily be concave down everywhere.  Consequently, linear interpolation of this function will systematically give erroneously low values.

\section{Results} \label{sec:results}
\subsection{Self Tests} \label{sec:test-self}
First, let's have a sanity check.  Using a cubic spline scheme with no smoothing to predict wavelengths for the returned, denoised line center values, $\lambda(x,y,m,n)$, should return identical values to the assigned true wavelength for each line within machine error.
This is what happens, for both ThArs and LFCs: look at the figure.

As an only slightly less silly sanity check, we can also use the constructed PCA basis to predict wavelengths for the measured line centers of an exposure that was included in the set of exposures used to construct the basis.  This interpolation should return identical PCA coefficients for that time.  Differences in wavelengths will arise from a combination of 1) the error inherent in measuring line centers and 2) interpolation error of lines to other x positions in the pixel direction.

Difference in denoised x-values to measured ones? (translate into rv shift?)
median and spread in errors in pixel (like 0.006) and RV (like 3.2 m/s?)
Histograms

\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
A better test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis and weights with PCA.  We can then use the results from that PCA to predict wavelengths for lines in the calibrations images that were left out.  We will call the used calibration images the ``training set," and the left out exposures the ``validation set."

We chose to use 90\% of available exposures to train the PCA, and leave the remaining 10\% for validation.  Exposures were sorted into either set at random.  This will test how well PCA coefficients are interpolated to other times outside of the training set and therefore how well line centers can be reconstructed at all times.  As before, differences in wavelength will also be drawn from how well line centers are measured and error in interpolating to those measured values.  This second source of error was independently quantified in the second test described in the previous section.

LFC looks good, ThAr is struggling a bit
Difference in denoised x-values to measured ones? (translate into rv shift?)
median and spread in errors in pixel (like 0.006) and RV (like 3.2 m/s?)
Histograms

\subsection{Cross Tests} \label{sec:test-cross}
Similar to the training and validation tests, we can also use one calibration source to construct the PCA basis and weights and the other to test the results.  For example, we can use LFC exposures to construct the wavelength solution, and then use it to predict wavelengths for measured line centers of ThAr lines, which will have their own assigned true wavelengths.  Likewise, we can use ThAr exposures to predict the wavelengths of LFC lines.  Note that errors in this test will also draw on the difference in quality of the two calibration sources being compared.

LFC predicts ThAr better than ThAr can predict itself.
ThAr predicting LFC is really messing up somehow, and I should figure out what it is (sorry!)

%\subsection{ThAr Stuff?  Or Include it in Previous Subsections?}

\section{Comparisons to Other Methods} \label{sec:comparisons}
I think we probably have to use the even/odd method to compare.

\subsection{Comparison to Parametric Model}
Previously, the \expres\ pipeline the $\theta_{n}$ from equation 1 comprises the
amplitudes of some 9th-ish-order polynomial in $x$ and $m$ for each $n$.
If so, they have
\begin{equation}
\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
\quad ,
\end{equation}
where the $c_{nij}$ are coefficients unique to exposure $n$ but smoothed by a third order polynomial over time.
Further, they were finding the coefficients $c_{nij}$ that
minimize an objective $Q$ that is something like the L2-norm:
\begin{equation}
Q = ||\lambda(x,y,m,n) - \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j||_2^2
\quad .
\end{equation}

Even in this simple context, where every calibration image $n$ is
treated as its own unique flower, there are improvements to be
made.
For one, the sums should not be from 0 to 9, but instead
\begin{equation}
\sum_{i=0}^9\sum_{j=0}^{9-i}
\quad ,
\end{equation}
because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.
For another, the objective function could be made soft to permit
catastrophic outliers without destroying the fit. (\lz{we did not do this})
We might recommend the iteratively reweighted least squares (IRLS).
This would make the fitting more robust.
For yet another, there are rescaling issues for the products $x^i\,m^j$ to
protect the fitting from near-singularities or bad conditioning of the
linear-algebra operators.

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/polyval2d.pdf}
\caption{}
\label{fig:polyValFit}
\end{figure}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/designMatrix.pdf}
\caption{}
\label{fig:dsnMFit}
\end{figure}

\subsection{Comparison to Non-Parametric, Non-Hierarchical Model}

\begin{figure}[h]
\centering
\includegraphics[width=0.9\textwidth]{Figures/noHierc.pdf}
\caption{}
\label{fig:intpFit}
\end{figure}

\section{Real Data Through to RVs} \label{sec:realdata}

\section{Discussion} \label{sec:discussion}

Generalization to etalon

Maybe some practical statements?  How many exposures are needed to implement this method at all.  How often do exposures have to be taken throughout a night (unless that's too dependent on instrument stability)

Some ideas about self-calibration. Once you know the space in which the
spectrograph lives, every single science frame gives you information about
where you are in that space; you don't need to look at the \lfc.

Chosing interpolations
monotonic vs. smooth (univariate spline stuff)

perfect degenercy between definition of "Centroid of line" and "wavelength solution"
cross correlation calls center of line something slightly different than what's happening when lines are being fit
may lead to systematic differences
in principle it would be preferable for the calibration source to look like a star

>>>>

\paragraph{Abstract:}
\name is a hierarchical, non-parametric framework for wavelength calibrating precision spectrographs.  There have been many recent hardware improvements in highly-stabilized, extreme-precision radial-velocity spectrographs.  In such stabilized spectrographs, the full optical system and detectors will only vary within a tiny range of configurations.  Additionally, within a given range of stability, every calibration image ever taken contains information relevant to every science exposure ever taken.  The development of laser-frequency combs or etalons allows the spectral range of an instrument to be covered with dense calibration points.  Here, we describe a calibration method---\name---that is designed for this new era of stabilized instruments with high-line-density calibrators.  We find that a high density of calibration lines allows for the freedom of a non-parametric wavelength solution that can match or adapt to any instrument or detector oddities.  We can also denoise the available calibration data by using it to determine not just the wavelength solution, but also the possible space of wavelength solutions, which is small for a stabilized instrument.  We demonstrate the success of this method with data from a Menlo Systems laser frequency comb taken by \expres, the Extreme Precision Spectrograph.  We find that \name\ returns more accurate wavelength predictions than classic polynomial models.
\section{Introduction}

Extreme precision radial-velocity (\eprv) programs have been
outrageously successful in finding and characterizing extra-solar
planets (CITE THINGS).
These programs typically make use of spectrographs with resolutions on
the order of $10^5$, which correspond to line widths on the order of
$3000\,\mps$.
Exoplanet science happens at $3\,\mps$ precisions, with current
instruments hoping to reach $0.1\,\mps$ precision (CITE THINGS).
This requires new spectrographs to be calibrated or
stabilized to better than $10^{-4}$ of a pixel (assuming that the
spectrograph is well sampled).
Some hardware--software systems seem to be reaching these
levels (CITE THINGS) .

Here we propose to simplify and improve calibration programs for
\eprv\ hardware systems with two very simple but novel ideas.
The first flows from the observation that calibration sources---which
include lamps, etalons, and laser-frequency combs (\lfc
s)---illuminate the spectrograph with very stable, very dense sets of lines; almost every
location in the spectrograph image plane is surrounded by nearby,
useful calibration lines.
This recommends a calibration methodology that is
\emph{non-parametric}:
If every point in the spectrograph detector is surrounded by nearby
calibration lines, the wavelength solution can, for example,  be made simply an interpolation of the calibration data.

We expand on the work of Milakovi\'{c}2020, who proposed constructing a wavelength solution as multiple, segmented polynomials across an order by proposing a wavelength solution that is completely non-parametric.
\lz{Should work in their paper more and praise it more?}

The density of lines removes the need to enforce any functional form for the wavelength solution (such as an eighth-order polynomial, for example).
Going non-parametric will improve calibration accuracy because
choosing a parametric form biases the calibration, especially when the chosen function is inappropriate (as, for example, polynomials will be at detector edges).

The second idea flows from the observation that contemporary
\eprv\ instruments are incredibly stable.
Temperature-controlled, fiber-fed pectrographs vary only slightly over
the night or season, and only along a small number of axes in what you
might call ``calibration space'', or the (very high dimensional) space
of all possible wavelength solutions.
That is, not only are the spectrographs stable, but they also have few accessible degrees of freedom.
This renders it inappropriate to fit each calibration exposure or calibrate
each science exposure independently.
Instead, all the calibration data (or all the data) should be used to
determine the space in which the instrument can and does vary.  Subsequent calibration work then need only determine where in the accessible part of calibration space the spectrograph existed.

This structure is \emph{hierarchical}: the calibration data are used
not just to determine the wavelength solution, but also to determine
the possible space of wavelength solutions.
In the statistics literature this concept is often described as
\emph{de-noising}.  
Every calibration exposure contains information about every other
exposure.  Thus every exposure can be improved (de-noised) with information coming from every other exposure.

The method we propose here---\name---embodies these ideas.
It is a non-parametric, hierarchical, data-driven model for the
wavelength solution.
By being non-parametric, it delivers enormous freedom to the
wavelength solution to match or adapt to any instrument or detector
oddities.
By being hierarchical, it restricts that freedom tremendously, but it
does so appropriately for the observed variations in the
spectrograph.
\name\ is designed for temperature-controlled, fiber-fed
spectrographs with good calibration sources, such as laser-frequency
combs, etalons, or good arc lamps.
We have in mind the \eprv\ instruments and \eprv\ science cases, but we expect \name\ to have applications for other kinds of
spectrographs in other contexts.  The motivating ideas behind this
project are true for almost every astronomical spectrograph.

\section{Method} \label{sec:method}
We operate on the two core ideas that the wavelength solution should be given enormous flexibility, but that it lives in a very low-dimensional space, where the degrees of freedom are set by the limited kinematics of the stabilized, spectrograph hardware.

We will pose the problem in the following way.  Given an exposure $n$, and order $m$, there is a relationship between
the two-dimensional $(x,y)$-position on the detector and the
wavelength $\lambda$
\begin{equation}
\lambda(x,y,m,n) = f(x,y,m;\theta_{n})
\quad ,
\end{equation}
where $\theta_{n}$ represents the parameters describing a given exposure.

A stabilized spectrograph should experience only low-degree variability, meaning the calibration of any image can be informed by the calibration of every other image.  To implement such a hierarchical model, we can use the calibration data themselves to develop a low-dimensional basis for expressing the calibration data.

If the space of all calibration possibilities is in fact $K$-dimensional (where K is a small integer, i.e 2 or 8 or thereabouts), and if the calibration variations are so small that we can linearize, then the function $f(x,y,m;\theta_{n})$ could be replaced with a tiny model
\begin{equation}
\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)
\quad ,
\end{equation}
where
$g_0(x,y,m)$ is the fiducial or mean or standard calibration of the spectrograph,
the $a_{nk}$ are scalar amplitudes,
and the $g_k(x,y,m)$ are basis functions expressing the ``directions'' in calibration space that the spectrograph can depart from the fiducial calibration.

The challenge is to learn these basis functions from the data and get the $K$ amplitudes, $a_{nk}$, for every exposure $n$.  There are many ways to discern these basis functions.  In this paper, we present a model using principal component analysis (PCA).  This is the correct thing to do in the limit where we have very high SNR, as is undeniably the case with calibration images.

\lz{Is this the place to mention options other than PCA? Or should we save that for the discussion?}

\subsection{Data} \label{sec:data}
We present results using data from \expres\, the EXtreme PRecison Spectrograph.  \expres\ is an environmentally-stabilized, fiber-fed, $R\sim137,000$, optical spectrograph (\lz{Jurgenson2016, Blackman2020}).  \expres\ has two different wavelength calibration sources, a ThAr lamp and a Menlo Systems laser frequency comb (LFC, e.g. \lz{Wilken2012, Molaro2013, Probst2014}).  ThAr exposures are taken at the beginning and end of each night.  LFC exposures are interspersed between science exposures every ~15-30 minutes.

The results and discussion presented here are based off of 1227 LFC exposures taken over two months months on 29 unique days.  LFC lines cover 40 echelle orders, which contain 19203 lines.  Though we primarily work with LFC data, we at times discuss applications to arc lamps.  These results use 78 ThAr exposures that were taken over the same range of time.  ThAr lines cover all 86 extracted orders of \expres\ and includes 5295 lines.

For the results presented in this paper, we work with the 1D, optimally extracted data.  This removes the y-dependence from equations 1 and 2.  Gaussian profiles are fit to each line in the 1D, extracted data.  The mean of each fitted Gaussian is taken to be the center of the line (see \lz{Petersburg2020} for more details).  While precisely defining and determining line positions is essential to precise wavelength calibration, we consider the task out of scope for this paper.  \name\ proceeds from a given list of orders, line wavelengths, and pixel positions of each line for each exposure.

\expres\ exposures are separated into different ``instrumental epochs," which correspond to atypical, hardware changes that affected the position of the echellogram on the detector, the shape of the instrumental PSF, or the calibration sources.  Because significant instrumental changes break the assumption of only low-dimensional variations in the spectrograph hardware, we treat each instrument epoch independently.

\subsection{Hierarchical De-Noising of Calibration Frames} \label{sec:denoising}
\lz{Concerns: 1) how to best phrase the problem using variables?, 2) how in-depth is too in-depth?}

Let $\{exp_n\}$ be the set of calibration exposures within an instrumental epoch.  These exposures will be used in constructing $g_0(x,y,m)$ and $g_k(x,y,m)$ in equation 2.  We first identify the full list of lines we can expect to find in a calibration image from this epoch.  A line is uniquely defined by a combination of order, $m$, and ``true" or theoretical wavelength assigned to that line, $\lambda$.  Let $\{l_{m,\lambda}\}$ be the complete list of unique lines found in $\{exp_n\}$.

Each line, $l_{m,\lambda}$, has a fitted line center, $x_{n,l}$, in pixel space, for each calibration exposure $exp_n$.  If this line is missing from an exposure (e.g. the fit failed due to noise, the line isn't in its usual order, etc.), we set $x_{n,l} = NaN$ for that line center in that exposure.  This allows us to construct a $N \times M$ matrix of line positions for each of $M$ lines for each of $N$ exposures.

Lines that are missing from greater than some percent of exposures are completely cut from the analysis.  Exposures that are missing greater than some percent of lines are also cut.  We use a threshold percentage of 50\% for both lines and exposures.  This largely affects the first lines on either end of an order and exposures with spuriously low signal.  Typically, about 0.2\% of lines and cut for LFC based calibration and 0.25\% of files are cut.

Missing line center measurements are replaced with denoised estimates using iterative PCA.  For each missing line center measurement, we initialize its value to the mean of the nearest $h$ exposures in time, where $h$ is typically $9$ exposures.  We then find a fiducial calibration of the spectrograph,  $g_0(x,y,m)$, which is set to be the mean line center value for each identified line over all exposures.  We subtract this fiducial calibration from each exposure's measured line centers and run PCA on the difference between the measured and mean line center for each exposure.  The resulting principal components are $g_k(x,y,m)$ in equation 2 while the amplitude of each principal component is $a_{nk}$.

Using those values with equation 2 and some small integer (e.g. 6) for $K$, we can reconstruct a denoised matrix of line centers for each exposure.  Missing line center measurements are replaced with values from this reconstruction.  We iterate until the values of the missing line centers change by less than 0.01\% of a pixel.  This convergence suggests that the PCA is no longer being swayed by outliers and is truly ``de-noised."   We have found this typically takes less than 30 iterations.

\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: $m$, order of each line; $\lambda$, assigned wavelength of each line; $\{x_{n,l}\} $, matrix of pixel line position of each line $l$ and for each exposure $n$\;
\While{change in missing or outlier line centers $>$ 0.01\%}{
	$g_0(x,y,m) = \overline{\{x_{n,l}\}}$\;
	find $U, \Sigma, V$ s.t. $U\Sigma V^* = (x_{l,n}-g_0(x,y,m))$\;
	let $a_{n,k} = U\cdot \Sigma$ and $V = g_k(x,y,m)$\;
	$\lambda(x,y,m,n) = g_0(x,y,m) + \sum_{k=1}^K a_{nk}\,g_k(x,y,m)$ for $K=6$\;
	$\{x_{n,l}\} = \lambda(x,y,m,n)$ where $\{x_{n,l}\}$ was initially $NaN$
	}
\caption{Hierarchical De-Noising}
\end{algorithm}

Following the de-noising process, we check for deviant line centers.  This is aimed at catching lines that have been mis-identified or poorly fit, resulting in erroneous line centers.  Outliers are identified via a simple sigma cut for each line.  Any line center that is reported at more than $3\sigma$ away from the mean line center value is relabeled a ``bad line.''  We then repeat the process of hierarchical de-noising to also replace the line centers of these bad lines. (\lz{Should probably add something about why the sigma cut works?})

\subsection{Non-Parametric Calibration} \label{sec:nonparam}
The process of hierarchical de-noising gives us $a_{nk}$, which captures the variability of the instrument.  We can interpolate $a_{nk}$ with respect to housekeeping data to determine line positions, $\lambda(x,y,m,n)$, for any state of the spectrograph.  In this paper, we show results where $a_{nk}$ is interpolated with respect to time, but it is possible to incorporate other information such as temperature, pressure, or telescope position.

Once we have the line center for every line in $\{l_{m,\lambda}\}$, each of which is associated with a known wavelength, we can interpolate  the known wavelengths over line centers to any position in x in a given order.  For example, interpolating onto every integer x will generate wavelengths for each pixel in an order.
\begin{algorithm}
\SetAlgoLined
\textbf{Inputs}: $g_0(x,y,m)$; $a_{n,k}$; $g_k(x,y,m)$; $t_n$, time of each exposure; $t'$, time we want wavelengths for\;
Interpolate $a_{n,k}$ with respect to time to determine $a_{n,k}'$ for time $t'$\;
Calculate pixel line positions for time $t'$ by evaluating\;
$\lambda(x,y,m,n)' = g_0(x,y,m) + \sum_{k=1}^K a_{nk}'\,g_k(x,y,m)$ for $K=6$\;
\For{each unique $m$}{
	interpolate ${l_n,m}$ onto pixels that we want wavelengths for\;
	}
\caption{Non-Parametric Wavelength Solution}
\end{algorithm}

\section{Tests}
\subsection{Self Tests} \label{sec:test-self}
For these so-called ``self tests", we use \name\ to  predict wavelengths for the measured line centers of a calibration exposure that was used to construct the basis.  The linear interpolation in time should return identical PCA coefficients for this time, as the calibration data was used in the PCA.

The predicted wavelengths for these measured line centers should match the assigned wavelengths for these calibration lines within the range of error of the measured line center.  Per the EXPRES pipeline, the LFC line centers used in this analysis are measured by fitting each LFC line to a Gaussian \lz{Peteresburg2020}.  The mean of the fitted Gaussian is set to be the line center.  The error in this fitted value is on average 2.1 $\mps$.  Additional error will arise from interpolation error of wavelengths to other pixel positions.

\subsection{Training and Validation Tests} \label{sec:test-trainNvalid}
A more full-scale test is to leave out some percentage of the calibration images from $\{exp_n\}$ while constructing the basis vectors and weights with PCA.  We can then use the resulting basis to predict wavelengths for lines in the calibrations images that were left out.  The used calibration images represent the ``training set," while the left out exposures make up the ``validation set."

For these tests, we used 90\% of available exposures to train the PCA, and left the remaining 10\% for validation.  Exposures were sorted into either set at random.  Contrary to the self tests, these results may also accrue errors from interpolating the principal component amplitudes with respect to time, i.e. we are now also testing how well we can re-construct line centers for all times.   As before, differences in wavelength will also be drawn from how well line centers are measured in an exposure and any error in interpolating to those measured values.  This second source of error was independently quantified in the second test described in the previous section.

\lz{I had some discussion of cross tests (i.e. using LFC to predict ThAr and ThAr to predict LFC, but I removed it since we decided to leave ThAr mostly out of this due to its data generally being worse}

\section{Results} \label{sec:results}

\begin{figure}[h]
\centering
\includegraphics[width=.5\textwidth]{Figures/lfcgood.png}
\caption{Residuals of different wavelength calibration tests.  Shown as a black outline is the distribution of residuals from the self test described in Section \ref{sec:test-self}, which incorporates errors in the fitted line centers and interpolating with respect to pixel.  Over-plotted in blue are the residuals from the training and validation test described in Section \ref{sec:test-trainNvalid}, which builds on the self tests by additionally incorporating possible errors from interpolating line positions through time.\
\lz{Considering adding a histogram of line center errors to show it is on par?  i.e. the limitation is the line fit.  What other information is needed?  i.e. the median/mean or spread?}}
\label{fig:testHists}
\end{figure}

Results from these tests are shown in Figure \ref{fig:testHists}.  The result of the self tests for LFC exposures is shown as a black outline.  This histogram combines residuals from all exposures.  The self test represents the limit of how well the line centers at any given time can be predicted, since here we are predicting the line centers for an exposure used in the construction of $a_{nk}$.

Overplotted in blue are the residuals from training the wavelength solution with only 90\% of LFC exposures and predicting the wavelengths for the remaining 10\% of exposures.  The two histograms match very closely, indicating that very little error is introduced when predicting denoised x values for times not used in the construction of the wavelength solution.

\section{Comparison to Classic, Parametric Model} \label{sec:comparisons}
Classically, pipelines employ polynomials to construct smooth wavelength solutions for each exposure.  For example, the \expres\ pipeline described in \lz{Petersburg2020} prescribed $\theta_{n}$ from equation 1 to be the amplitudes of a 2D, 9th-order polynomial in $x$ and $m$ for each $n$, making the equivalent to equation 2 the following:
\begin{equation}
\lambda(x,y,m,n) = \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j + \mathrm{noise}
\quad ,
\end{equation}
where $c_{nij}$ are coefficients unique to exposure $n$.  These coefficients are then smoothed by a third-order polynomial over time.  This third-order polynomial would be evaluated at the time of non-calibration exposures to re-construct a 2D, 9th-order polynomial wavelength solution for that exposure.
Coefficients, $c_{nij}$, were found that minimize an objective $Q$ that is the L2-norm:
\begin{equation}
Q = ||\lambda(x,y,m,n) - \sum_{i=0}^9\sum_{j=0}^9 c_{nij}\, x^i\,m^j||_2^2
\quad .
\end{equation}

Even in this simple context, where every calibration image $n$ is treated as its own unique flower, there are improvements to be made.
For one, the sums should not be from 0 to 9, but instead
\begin{equation}
\sum_{i=0}^9\sum_{j=0}^{9-i}
\quad ,
\end{equation}
because that is the definition of 9th order.  We find that fitting to 8th order is sufficient, faster, and leaves less residual pattern than the original 9th degree fit.

For another, the objective function could be made soft to permit catastrophic outliers without destroying the fit. (\lz{we did not do this}).  We might recommend the iteratively reweighted least squares (IRLS).  This would make the fitting more robust.

For yet another, there are rescaling issues for the products $x^i\,m^j$ to protect the fitting from near-singularities or bad conditioning of the linear-algebra operators.

\subsection{Comparisons}
To test the parametric vs. the non-parametric method, we took an LFC exposure and separated the even and odd lines.  We then used the even lines to predict wavelengths for the odd lines and vice versa.

For example, for the parametric model, we fit a 2D, 9th-degree polynomial to the odd lines.  We then evaluated this polynomial at the positions of the even lines and compared those returned wavelengths to the assigned wavelength of each line.  We then reversed the process by fitting a polynomial to the even lines and evaluating that polynomial at the positions of the odd lines.  Differences between the polynomial given wavelengths and the assigned wavelengths of each line are shown in Figure \ref{fig:parametricTest}, left.
